\documentclass{article}

\title{Some suggestions on Perl Evangelism}
\author{Stuart Cooper\\stuart.cooper@gmail.com\\Sydney Perl User Group}
\date{August 20, 2025}


\begin{document}

\maketitle

This is a rehash of a talk I gave to the Sydney Perl User Group in
2015; almost exactly ten years ago. I look at some ways to
promote Perl to co-workers; who might be intelligent sysadmins,
Python/Ruby/Go Developers or other I.T workers. People
who spend a lot of time at the Linux command line. I finish with some
stories of introducting Perl and my own path to Perl.

\begin{itemize}
  \item Religious Evangelism: Don't preach, be a good example that
    people want to follow. Be patient, be nice.
  \item Make yourself known as The Perl Guy.
  \item Play to Perl's strengths. Regular Expressions. Glue
    language. Rapid development. Quick protoyping.
  \item Make sure you have the Perl that you want; both on your
    development machine and on your company's servers (nodes if you
    like). You're probably using Puppet or Ansible to keep them
    conformed so if you need a change across all machines, do it
    through approved mechanisms. You might need to package your
    modules and code as .rpm or .deb files; so learn how to do that. A
    modern program that's pretty exciting and good at doing this is
    the ``effing package manager'', fpm. It's written in ruby and
    might get me properly interested in ruby someday.
  \item \emph{Make sure that \emph{perldoc} is installed properly and that
  you can run it as root.} If you're demoing stuff at the shell on some
    random server and you reach for perldoc and it isn't there you're
    going to look pretty stupid and your co-worker will lose interest
    pretty quickly.
  \item Know handy command line switches and know some good Perl one
    liners. In particular know how to do common sed jobs in Perl and
    common awk jobs in Perl. Know how to do \verb!awk '{print $3}'!
    in Perl. When a guy showed me this I said ``well that's easy to do
    in awk'' and he simply said ``I don't need to know awk''.
  \item Avoid jargon, in-jokes and arcana that are fun to Perl people
    but forbidding and exclusionary to newcomers. Notice that in both
    my talks tonight I've said Sydney Perl User Group, not
    Sydney.pm. Sydney.pm is an injoke among Perl people, a Sydney Perl
    User Group is a User Group for Sydney Perl users cf. SLUG.
\end{itemize}

\section{Two Perl introductions}

You never know what helping a friend or colleague learn Perl or
install Perl will lead to. I worked for Cisco at the height of the
original Dot Com boom and was let go at the Dot Com Crash. Early on a
Vietnamese/Aussie young guy named Vietlong asked for help installing
Perl and writing a simple program, under Windows. I hadn't written Perl
on Windows before but the demonstration program I wrote on my Linux
laptop, a 12-line program that read and updated a file continuously
while it was running, worked first time on Vietlong's laptop with no
changes required. Portability!!

Vietlong didn't perservere with Perl and ended up writing a Java-clock
which looked nice and was used in various groups across Australian
Cisco.

The second guy to ask me for Perl help, three months later, was an older
Aussie guy named Adam. Again he wanted Perl installed on Windows,
which I now didn't even have to google for help on having done it for
Vietlong earlier. Adam didn't need any demo programs, once Perl was
properly installed on his Windows laptop he sent me on my way. We were
both in the Australian office but I was attached to a US Perl team and
most of the Aussies in my area were working with Java.

I bugged Adam for a week after; to make sure he was going OK
with Perl and wasn't giving up on it quickly like Vietlong did. He
brushed me off, he didn't want any more help from me. I found him
a bit rude though probably he was just very focussed.

Adam turned out OK at Perl. He was better than OK, he was ADAMK.

You never know what helping a colleague use Perl will lead to.

\subsection{SLUG installfests}

The best example of user group help I've personally seen are the
Sydney Linux Users Group Installfest days, in the early 2000s. They
had some rooms at Uni of Technology Sydney booked all Saturday and
people could come with their laptops (or even desktops in those days)
and SLUG volunteers would help install Linux on their machines.

The volunteers all had cool black T-shirts; I was a volunteer for an
installfest in 2000 and kept the cool t-shirt for years afterwards.
I was more useful for ``Now you've got Linux installed here's some
cool Day 1 things you can do with it'' than tricky installation
problems; also I was helping Redhat installations and the smartest
guys were the Debian guys and they monopolised as many installations
as they could!

2000 was a great Linux year for me. I attempted to name my son Linux,
but I never heard back from the email I sent to Linus Torvalds so I
erred on the safe side and went with the name Linus. On the morning of
Linus's birth; with baby and Mum safe in hospital, I went to
Everything Linux's shop in Five Dock and bought the book of
More Programming Pearls which I've had for as long as my son.

So with Perl, the likelihood is that your Linux colleague has Perl on
his desktop and systems and could benefit from a Day 1 demonstration
of some Perl programs and features. Try not to overwhelm your
colleague and you're not trying to show off your mad skillz and
superiority. See what your friend is interested in and tailor your
material so it's immediately useful to him.

\section{How I came to Perl}

After my University days (1989-1992) in Newcastle NSW, I got a first
job in Sydney for Unidata, a database running across all the UNIXes
and a migration path to UNIX systems for users of old systems called
Pick and PRIME. An important part of Unidata's IP were conversion
tools; to transform Pick Basic and PRIME Basic into Unidata's flavour
of Basic, UniBasic.

These tools were written in the most powerful data transforming tools
available in UNIX pre-Perl; lex and yacc. I was decently good at lex
at Uni and became much better at it at Unidata; sadly yacc always
terrified me and I never took the trouble to understand it.

There were over sixty versions of UNIX that Unidata ran on. The most
important were SunOS (later Solaris), HP/UX, IBM/AIX and SCO UNIX for
the smallest customers. Two more boutique UNIXes I spent a lot of time
with were Silicon Graphics UNIX and Data General UNIX, DG/UX, which
had Emacs, gcc and a bunch of goodies pre-installed.

I fell wildly in love with the O'Reilly books; particularly UNIX in a
Nutshell SVR4 edition. It was the best \$15 I'd ever spent; I still
have my copy and it's been used so much that the plastic film has been
scraped off the front cover through years of me touching the
book. One year things were pretty quiet over Christmas and my wise
boss encouraged me to learn Emacs from the chapter in UNIX in a
Nutshell and any other resources I could find. Three days later, I was
a happy Emacs user and remain just as happy today.

There was an O'Reilly book that was waaaaay overpriced, at \$120. Lex
and Yacc was pretty steep at \$60 but \$120 for a mega-book was too
much for me to spend on a book. My boss had it and occasionally if I
was good or I had a justified need to I could use his copy which he
kept in the office. That book was UNIX Power Tools; not with a
cutesy animal on the cover like a Tarsier or a Lex Bird and Yacc Bird,
but a power drill. It collected all the text processing advice from
other O'Reilly books and also had a lot of Perl content; which was
starting to be a hot new language in the UNIX world.

\subsection{My first Perl program}

I didn't write any Perl for Unidata but I did at my
first contract job, for Commonwealth Bank, in 1996. On AT\&T UNIX, I
was working in some directories with big files in them and my eyes got
tired of looking at ls -l output and figuring if a file of 17435346
bytes was 17 million or 1.7 million or 174 million bytes. I wanted
placeholder commas in the output, so the filesize would be reported
as 17,435,346.

Coming from a C and lex background and not so strong in shell; the
Perl program was verbose and not idiomatic Perl at all but it worked a
charm. Here's a quick 2025 version of it:

\begin{verbatim}
#!/usr/bin/perl
# pls: perl ls; run ls -l but stick placeholding commas in
# filesizes 5 digits or more

use strict;
use warnings;

my (@lslines) = `ls -l`;
foreach (@lslines) {
  chomp;
  s/(\d{5,})/addcomma($1)/e;
  print "$_\n";
}

sub addcomma {
  my ($n) = @_;
  my $r = q{};
  my $l = length($n);
  for ( 1 .. $l) {
    $r = substr( $n, $l-$_, 1 ) . $r;
    if ( $_ % 3 == 0 ) { $r = ',' . $r; }
  }
  $r = substr($r,1) if (substr($r,0,1) eq ',');
  return $r;
}
\end{verbatim}

Here's pls in action in my dwhelper directory
\begin{verbatim}
/home/scooper/dwhelper: ls -l | tail -2
-rw-rw-r-- 1 scooper scooper   27142819 Mar 29  2024 winx_cox_plate2.mp4
-rw-rw-r-- 1 scooper scooper    8711631 Dec  9  2024 zero_would_be_nice.mp4
/home/scooper/dwhelper: pls | tail -2
-rw-rw-r-- 1 scooper scooper   27,142,819 Mar 29  2024 winx_cox_plate2.mp4
-rw-rw-r-- 1 scooper scooper    8,711,631 Dec  9  2024 zero_would_be_nice.mp4
\end{verbatim}

\section{A Perl Revival?}

I'm a very nostalgic person and it might seem that this talk is a call
for computing to return to pre-web days of all the action being at the
command line. That isn't going to happen but doing jobs in Perl on
your own systems is a fantastic use case for Perl. My default thinking
for a personal project is ``I can do that in Perl''.

Languages come and go in popularity. PHP was in the doldrums for years
then enjoyed a comeback. Similar for Javascript. LISP was boutique
when I was at Uni but is coming back in various guises. Ruby on Rails
seems gone but Ruby itself is gaining mindshare.

I think Perl gets some very unfair treatment in discussions of
everyone who hasn't used it seriously. Perl is criticised for looking
ugly while PHP code can be as hideous as it likes, so long as it seems
to work on most browsers. I once wrote some very clear Perl for a
Nagios plugin to check usage of a datastore, at a non-Perl shop, and
got pilloried for using Perl, even though the example in the
datastore's API documentation was in Perl. The same company was using
complicated bash scripts and were dependent on the external program
shellcheck.net to try and save them from their own coding.

An excellent answer to the old slur that Perl is ``just a CGI
language'' is to have a bunch of programs on your system that use Perl
in many spaces: algorithms, networking, database access, cron jobs.

Perhaps there's a ``Killer App'' about to be released in Perl that
will reinvigorate interest in the language. Bitcoin, AI, something
like that. When that happens, existing Perl people will be well
positioned to ride the cusp of the New Wave, and install Perl on
Windows for the next generation of ADAMKs.

I was surprised a few years ago that Microsoft came up with Powershell
and didn't simply rejig or piggyback off Perl as their hot programming
language. If Windows had made Perl important, and commercial Linux
jobs had no need for Perl, I would have changed the habits of a
lifetime and switched to Windows, such is my love of Perl.

My 2015 version of this talk grew out of reading the first few
chapters of Tim Maher's book, ``Minimal Perl''. I recommend this book
as a kind of ``Simple Perl Best Practices''; if you code according to
Minimal Perl style your code will be accessible to smart Linux people
without a Perl background. I don't think you're giving up too much in
power and fun by doing this. You might miss perl golf and some of
the dusty corners of the language, but you make up for it by being
welcoming and inclusive.

A good way of doing some perl regularly is with the Perl Weekly
Challenge. Many of the solvers can't help pasting their solutions to
each week's challenge on Facebook.

% Backpage 6 with References on it: KISS with {description}
\section*{References and Resources}

\begin{description}

\item[Minimal Perl. Tim Maher. Manning,2007] An excellent book about
  using a clearly understood subset of Perl to welcome new users to
  Perl based on their existing experience. Explains Perl to shell
  folk, sed folk, awk fold and C folk. This book was the inspiration
  for my original Perl Evangelism talk in 2015.

\item[git@github.com:briandfoy/PerlPowerTools.git] The Perl Power
  Tools Project; an implementation in Perl of common Linux commands by
  brian d foy.

\item[Unix Power Tools. Peek/O'Reilly O'Reilly,1993] A
  major book of the early 1990s promoting Perl to UNIX users.
  The book included Tom Christiansen's ``C Shell Programming
  Considered  Harmful'' which helpfully put me off the C Shell
  forever.

\item[UNIX in a Nutshell SVR4 Edition Gilly. O'Reilly,1992] The best
  single-volume book on UNIX shells, sed and awk, emacs and vi, troff
  and groff and lots of other subjects. Everything UNIX before Perl,
  and just about everything a UNIX user (not a sysadmin) needs to know
  to navigate UNIX. Later editions and the book's godchild Linux in a
  Nutshell contain Perl content. A lot of ``Using Linux'' doorstop
  books copied the format of UNIX in a Nutshell; I didn't need these
  books because I was old enough to have studied UNIX in a Nutshell
  but some that I liked the look of were from an author named Mark
  Sobell.

\item[The UNIX Programming Environment. Kernighan \& Pike. Prentice
  Hall,1984] An early catch-all book which is about the same size and
  complexity   of Kernighan \& Ritchie The C Programming Language,
  that covers shell, C Programming, UNIX system calls, lex and
  yacc. Like K\&R C it assumes you know what you're doing and
  you have to be pretty scientific to read this book. The same author
  combination wrote The Practice of Programming, Addison Wesley,
  1999.

\item[The New Hacker's Dictionary. Eric S. Raymond. MIT Press,1991]
  Another classic book that we had at the Sydney Unidata office. We
  read it cover to cover and got a good sense of 1990s US Hacker
  culture from this fantastic book. It's the printed version of an
  online resource called the JARGON file, so if you don't want the
  printed book you can get the same information from the Jargon
  file.

% Effective Perl Programming? Lex and Yacc?
\end{description}

\end{document}
